Reviving the Lost Art of \\Combinator Graph Reduction

### 2023-01-29 experiments/superg/doc/paper/paper.tex
Insert refs for Kiselyov.

\subsection{Lazy weakening}
Adding lazy weakening requires that we track which items in the context are ignored.
We change our context representation from an integer to a list of boolean values.
In the new compilation functions below, we use a shorthand $t$ for the value \textit{true} and $f$ for \textit{false}.
We always match the last element of lists and represent the remainder of the list with $\Gamma$.

\begin{equation*}
    \begin{array}{l l l l}
        \lazy & z         & = & t \models I                                                             \\
        \lazy & s \ e     & = & \Gamma,f \models c                                                      \\
              &           &   & \text{where } (\Gamma \models c) := \lazy e                             \\
        \\
        \lazy & \lambda e & = & \emptyset \models K \; c                                                \\
              &           &   & \text{where } (\emptyset \models c) := \lazy e                          \\
        \lazy & \lambda e & = & (\emptyset \models K) \coprod (\Gamma \models c)                        \\
              &           &   & \text{where } (\Gamma,f \models c) := \lazy e                           \\
        \lazy & \lambda e & = & \Gamma \models K \; c                                                   \\
              &           &   & \text{where } (\Gamma,t \models c) := \lazy e                           \\
        \\
        \lazy & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)) \\
              &           &   & \text{where } (\Gamma_1 \models c_1) := \lazy e_1                       \\
              &           &   & \text{where } (\Gamma_2 \models c_2) := \lazy e_2                       \\
              &           &   & \text{where } \Gamma := \Gamma_1 \sqcup \Gamma_2                        \\
    \end{array}
\end{equation*}

We have introduced yet another case for $\lambda e$, this time to distinguish between $t$ or $f$ as the last element in the context of the inner expression.
There is also the new function $\sqcup$, which we need to implement:

\begin{equation*}
    \begin{array}{l l l l l}
        \Gamma_1    & \sqcup & \emptyset   & = & \Gamma_1                     \\
        \emptyset   & \sqcup & \Gamma_2    & = & \Gamma_2                     \\
        \Gamma_1,t  & \sqcup & \Gamma_2,\_ & = & (\Gamma_1 \sqcup \Gamma_2),t \\
        \Gamma_1,\_ & \sqcup & \Gamma_2,t  & = & (\Gamma_1 \sqcup \Gamma_2),t \\
        \Gamma_1,f  & \sqcup & \Gamma_2,f  & = & (\Gamma_1 \sqcup \Gamma_2),f \\
    \end{array}
\end{equation*}

@@

Our semantic function now has to handle more cases, and thus is somewhat larger:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        (\emptyset \models c_1)  & \coprod & (\emptyset \models c_2)  & = & c_1 \ c_2                                                \\
        \\
        (\emptyset \models c_1)  & \coprod & (\Gamma_2,t \models c_2) & = & (\emptyset \models B c_1) \coprod (\Gamma_2 \models c_2) \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\emptyset \models C C c_2) \coprod (\Gamma \models c_1) \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (s \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } s := (\emptyset \models S )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        % The new additions
        (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)    \\
        \\
        (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (b \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } b := (\emptyset \models B )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models (c \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } c := (\emptyset \models C )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        (\emptyset \models c_1)  & \coprod & (\Gamma_2,f \models c_2) & = & (\emptyset \models c_1) \coprod (\Gamma_2 \models c_2)   \\
        \\
        (\Gamma_1,f \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\Gamma_1 \models c_1) \coprod (\emptyset \models c_2)   \\
    \end{array}
\end{equation*}

@@

\subsection{Eta optimization}
Kiselyov covers eta optimization as a change to his OCaml code rather than typing rules.
Our recursive functions are derived from the typing rules, so these changes are not as straightforward to incorporate.
%Previously Kiselyov's representation consisted of elements $C$, $N$ and $W$, representing closed expressions, a used context element, and an ignored context element respectively.
The introduction of the $V$ element means that a context may now be associated with either a closed expression or the top variable (we reuse $V$ as a marker for this).
Our implementation uses a tagged union to define a type that can be either an expression or the top variable.
For the compilation function, we only need to change the definition of $\lazyeta \ z$ and add an extra case for $\lazyeta \ \lambda e$:

\begin{equation*}
    \begin{array}{l l l l}
        \lazyeta & z         & = & \emptyset \models V                                                          \\
        \lazyeta & s \ e     & = & \Gamma,f \models c                                                           \\
                 &           &   & \text{where } (\Gamma \models c) := \lazyeta e                               \\
        \\
        \lazyeta & \lambda e & = & \emptyset \models I                                                          \\
                 &           &   & \text{where } (\emptyset \models V) := \lazyeta e                            \\
        \lazyeta & \lambda e & = & \emptyset \models K \; c                                                     \\
                 &           &   & \text{where } (\emptyset \models c) := \lazyeta e                            \\
        \lazyeta & \lambda e & = & (\emptyset \models K) \coprod_\eta (\Gamma \models c)                        \\
                 &           &   & \text{where } (\Gamma,f \models c) := \lazyeta e                             \\
        \lazyeta & \lambda e & = & \Gamma \models K \; c                                                        \\
                 &           &   & \text{where } (\Gamma,t \models c) := \lazyeta e                             \\
        \\
        \lazyeta & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod_\eta (\Gamma_2 \models c_2)) \\
                 &           &   & \text{where } (\Gamma_1 \models c_1) := \lazyeta e_1                         \\
                 &           &   & \text{where } (\Gamma_2 \models c_2) := \lazyeta e_2                         \\
                 &           &   & \text{where } \Gamma := \Gamma_1 \sqcup_\eta \Gamma_2                        \\
    \end{array}
\end{equation*}

Our context merge function $\sqcup$ now also depends on the compiled expression, since we have a special case for $V$:

\begin{equation*}
    \begin{array}{l l l l l}
        (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\emptyset \models V)     & = & \Gamma_1,t                        \\
        (\emptyset \models V)     & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & \Gamma_2,t                        \\
        (\emptyset \models c_1)   & \sqcup_\eta & (\emptyset \models V)     & = & t                                 \\
        (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models c_2)   & = & t                                 \\
        (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models V)     & = & \mathbf{impossible}               \\
        \\
        % Original ones
        (\Gamma_1 \models c_1)    & \sqcup_\eta & (\emptyset \models c_2)   & = & \Gamma_1                          \\
        (\emptyset \models c_1)   & \sqcup_\eta & (\Gamma_2 \models c_2)    & = & \Gamma_2                          \\
        (\Gamma_1,t \models c_1)  & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
        (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\Gamma_2,t \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
        (\Gamma_1,f \models c_1)  & \sqcup_\eta & (\Gamma_2,f \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),f \\
    \end{array}
\end{equation*}

The $V$ value effectively encodes an additional implicit used element in the context.
We also have some new cases for the semantic function:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        % Added for eta optimization
        (\Gamma_1,f \models c_1) & \coprod_\eta & (\emptyset \models V)    & = & c_1                                                        \\
        (\emptyset \models V)    & \coprod_\eta & (\Gamma_2,f \models c_2) & = & (\emptyset \models CI) \coprod_\eta (\Gamma_2 \models c_2) \\

        (\Gamma_1,t \models c_1) & \coprod_\eta & (\emptyset \models V)    & = & (\Gamma_1 \models (s \coprod_\eta l)) \coprod_\eta i       \\
                                 &              &                          &   & \text{where } l := (\Gamma_1 \models c_1 )                 \\
                                 &              &                          &   & \text{where } s := (\emptyset \models S )                  \\
                                 &              &                          &   & \text{where } i := (\emptyset \models I )                  \\
        (\emptyset \models V)    & \coprod_\eta & (\Gamma_2,t \models c_2) & = & (\emptyset \models SI) \coprod_\eta (\Gamma_2 \models c_2) \\
        (\emptyset \models c_1)  & \coprod_\eta & (\emptyset \models V)    & = & c_1                                                        \\
        (\emptyset \models V)    & \coprod_\eta & (\emptyset \models c_2)  & = & CI c_2                                                     \\
        (\emptyset \models V)    & \coprod_\eta & (\emptyset \models V)    & = & \mathbf{impossible}                                        \\
        \\

        (\Gamma_1 \models c_1)   & \coprod_\eta & (\Gamma_2 \models)       & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models)          \\
    \end{array}
\end{equation*}

@@

\subsection{Linear algorithm}
The implementation of the linear algorithm is much simpler because it is based on the $\strict$ algorithm.
The compilation function is identical save for a different semantic function $\coprod_{\text{linear}}$:

\begin{equation*}
    \begin{array}{l l l l}
        \linear & z         & = & 1 \models I                                                              \\
        \linear & s \ e     & = & n+1 \models (0 \models K) \coprod (n \models c)                          \\
                &           &   & \text{where } (n \models c) := \linear e                                 \\
        \\
        \linear & \lambda e & = & 0 \models K \; c                                                         \\
                &           &   & \text{where } (0 \models c) := \linear e                                 \\
        \linear & \lambda e & = & (n-1) \models K \; c                                                     \\
                &           &   & \text{where } (n \models c) := \linear e                                 \\
        \\
        \linear & e_1 \ e_2 & = & (n \models ((n_1 \models c_1) \coprod_{\text{linear}} (n_2 \models c_2)) \\
                &           &   & \text{where } (n_1 \models c_1) := \linear e_1                           \\
                &           &   & \text{where } (n_2 \models c_2) := \linear e_2                           \\
                &           &   & \text{where } n := \max n_1 \, n_2                                       \\
    \end{array}
\end{equation*}

below we give the corresponding semantic function, which is no longer recursive, making this algorithm linear:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        (0 \models c_1)   & \coprod & (0 \models c_2)   & = & c_1 \ c_2                               \\
        \\
        (0 \models c_1)   & \coprod & (n \models c_2)   & = & B_n \ c_1 \ c_2                         \\
        \\
        (n \models c_1)   & \coprod & (0 \models c_2)   & = & C_n \ c_1 \ c_2                         \\
        \\
        (n \models c_1)   & \coprod & (n \models c_2)   & = & S_n \ c_1 \ c_2                         \\
        \\
        (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & B_{n_2-n_1} \ (S_{n_1} \ c_1) \ c_2     \\
                          &         &                   &   & \text{where } n_1 < n_2                 \\
        (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & C_{n_1-n_2} \ (B_{n_1-n_2} \ c_1) \ c_2 \\
                          &         &                   &   & \text{where } n_1 > n_2                 \\
    \end{array}
\end{equation*}

@@

\begin{figure*}
    \begin{equation*}
        \begin{array}{l l l l}
            \lazy & z         & = & t \models I                                                             \\
            \lazy & s \ e     & = & \Gamma,f \models c                                                      \\
                  &           &   & \text{where } (\Gamma \models c) := \lazy e                             \\
            \\
            \lazy & \lambda e & = & \emptyset \models K \; c                                                \\
                  &           &   & \text{where } (\emptyset \models c) := \lazy e                          \\
            \lazy & \lambda e & = & (\emptyset \models K) \coprod (\Gamma \models c)                        \\
                  &           &   & \text{where } (\Gamma,f \models c) := \lazy e                           \\
            \lazy & \lambda e & = & \Gamma \models K \; c                                                   \\
                  &           &   & \text{where } (\Gamma,t \models c) := \lazy e                           \\
            \\
            \lazy & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)) \\
                  &           &   & \text{where } (\Gamma_1 \models c_1) := \lazy e_1                       \\
                  &           &   & \text{where } (\Gamma_2 \models c_2) := \lazy e_2                       \\
                  &           &   & \text{where } \Gamma := \Gamma_1 \sqcup \Gamma_2                        \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Compilation function with Lazy Weakening.}
    \label{fig:lazy}
\end{figure*}

\begin{figure*}
    \begin{equation*}
        \begin{array}{l l l l l}
            \Gamma_1    & \sqcup & \emptyset   & = & \Gamma_1                     \\
            \emptyset   & \sqcup & \Gamma_2    & = & \Gamma_2                     \\
            \Gamma_1,t  & \sqcup & \Gamma_2,\_ & = & (\Gamma_1 \sqcup \Gamma_2),t \\
            \Gamma_1,\_ & \sqcup & \Gamma_2,t  & = & (\Gamma_1 \sqcup \Gamma_2),t \\
            \Gamma_1,f  & \sqcup & \Gamma_2,f  & = & (\Gamma_1 \sqcup \Gamma_2),f \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Semantic function with Lazy Weakening.}
    \label{fig:lazy-sem}
\end{figure*}

\begin{figure*}
    \begin{equation*}
        \begin{array}{l l l l}
            \lazyeta & z         & = & \emptyset \models V                                                          \\
            \lazyeta & s \ e     & = & \Gamma,f \models c                                                           \\
                     &           &   & \text{where } (\Gamma \models c) := \lazyeta e                               \\
            \\
            \lazyeta & \lambda e & = & \emptyset \models I                                                          \\
                     &           &   & \text{where } (\emptyset \models V) := \lazyeta e                            \\
            \lazyeta & \lambda e & = & \emptyset \models K \; c                                                     \\
                     &           &   & \text{where } (\emptyset \models c) := \lazyeta e                            \\
            \lazyeta & \lambda e & = & (\emptyset \models K) \coprod_\eta (\Gamma \models c)                        \\
                     &           &   & \text{where } (\Gamma,f \models c) := \lazyeta e                             \\
            \lazyeta & \lambda e & = & \Gamma \models K \; c                                                        \\
                     &           &   & \text{where } (\Gamma,t \models c) := \lazyeta e                             \\
            \\
            \lazyeta & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod_\eta (\Gamma_2 \models c_2)) \\
                     &           &   & \text{where } (\Gamma_1 \models c_1) := \lazyeta e_1                         \\
                     &           &   & \text{where } (\Gamma_2 \models c_2) := \lazyeta e_2                         \\
                     &           &   & \text{where } \Gamma := \Gamma_1 \sqcup_\eta \Gamma_2                        \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Compilation function with Eta optimization.}
    \label{fig:lazyeta}
\end{figure*}

\begin{figure*}

    \begin{equation*}
        \begin{array}{l l l l l}
            (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\emptyset \models V)     & = & \Gamma_1,t                        \\
            (\emptyset \models V)     & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & \Gamma_2,t                        \\
            (\emptyset \models c_1)   & \sqcup_\eta & (\emptyset \models V)     & = & t                                 \\
            (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models c_2)   & = & t                                 \\
            (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models V)     & = & \mathbf{impossible}               \\
            \\
            % Original ones
            (\Gamma_1 \models c_1)    & \sqcup_\eta & (\emptyset \models c_2)   & = & \Gamma_1                          \\
            (\emptyset \models c_1)   & \sqcup_\eta & (\Gamma_2 \models c_2)    & = & \Gamma_2                          \\
            (\Gamma_1,t \models c_1)  & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
            (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\Gamma_2,t \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
            (\Gamma_1,f \models c_1)  & \sqcup_\eta & (\Gamma_2,f \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),f \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Context merge function for Eta optimization.}
    \label{fig:lazyeta-ctx}
\end{figure*}

\begin{figure*}
    \begin{equation*}
        \arraycolsep=1.4pt
        \begin{array}{l l l l l}
            % Added for eta optimization
            (\Gamma_1,f \models c_1) & \coprod_\eta & (\emptyset \models V)    & = & c_1                                                        \\
            (\emptyset \models V)    & \coprod_\eta & (\Gamma_2,f \models c_2) & = & (\emptyset \models CI) \coprod_\eta (\Gamma_2 \models c_2) \\

@@

            (\Gamma_1 \models c_1)   & \coprod_\eta & (\Gamma_2 \models)       & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models)          \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Semantic function with Eta optimization.}
    \label{fig:lazyeta-sem}
\end{figure*}

\begin{figure*}
    \begin{equation*}
        \begin{array}{l l l l}
            \linear & z         & = & 1 \models I                                                              \\
            \linear & s \ e     & = & n+1 \models (0 \models K) \coprod (n \models c)                          \\
                    &           &   & \text{where } (n \models c) := \linear e                                 \\
            \\
            \linear & \lambda e & = & 0 \models K \; c                                                         \\
                    &           &   & \text{where } (0 \models c) := \linear e                                 \\
            \linear & \lambda e & = & (n-1) \models K \; c                                                     \\
                    &           &   & \text{where } (n \models c) := \linear e                                 \\
            \\
            \linear & e_1 \ e_2 & = & (n \models ((n_1 \models c_1) \coprod_{\text{linear}} (n_2 \models c_2)) \\
                    &           &   & \text{where } (n_1 \models c_1) := \linear e_1                           \\
                    &           &   & \text{where } (n_2 \models c_2) := \linear e_2                           \\
                    &           &   & \text{where } n := \max n_1 \, n_2                                       \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Compilation function for Linear algorithm}
    \label{fig:linear}
\end{figure*}

TODO

\begin{figure*}
    \begin{equation*}
        \arraycolsep=1.4pt
        \begin{array}{l l l l l}
            (0 \models c_1)   & \coprod & (0 \models c_2)   & = & c_1 \ c_2                               \\
            \\
            (0 \models c_1)   & \coprod & (n \models c_2)   & = & B_n \ c_1 \ c_2                         \\
            \\
            (n \models c_1)   & \coprod & (0 \models c_2)   & = & C_n \ c_1 \ c_2                         \\
            \\
            (n \models c_1)   & \coprod & (n \models c_2)   & = & S_n \ c_1 \ c_2                         \\
            \\
            (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & B_{n_2-n_1} \ (S_{n_1} \ c_1) \ c_2     \\
                              &         &                   &   & \text{where } n_1 < n_2                 \\
            (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & C_{n_1-n_2} \ (B_{n_1-n_2} \ c_1) \ c_2 \\
                              &         &                   &   & \text{where } n_1 > n_2                 \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Semantic function for Linear algorithm}
    \label{fig:linear-sem}
\end{figure*}

>>>>

\subsection{Lazy weakening}
Adding lazy weakening requires that we track which items in the context are ignored.
We change our context representation from an integer to a list of boolean values.
In the new compilation functions below, we use a shorthand $t$ for the value \textit{true} and $f$ for \textit{false}.
We always match the last element of lists and represent the remainder of the list with $\Gamma$.
Appendix~\ref{app:kiselyov} contains the new compilation function $\lazy$ (Figure~\ref{fig:lazy}).

We have introduced yet another case for $\lambda e$, this time to distinguish between $t$ or $f$ as the last element in the context of the inner expression.
There is also the new function $\sqcup$, whose definition is given in Figure~\ref{fig:lazy-ctx}.

@@

Our semantic function now has to handle more cases, and thus is somewhat larger.
See Figure~\ref{fig:lazy-sem} for the full definition.

@@

\subsection{Eta optimization}
Kiselyov covers eta optimization as a change to his OCaml code rather than typing rules.
Our recursive functions are derived from the typing rules, so these changes are not as straightforward to incorporate.
%Previously Kiselyov's representation consisted of elements $C$, $N$ and $W$, representing closed expressions, a used context element, and an ignored context element respectively.
The introduction of the $V$ element means that a context may now be associated with either a closed expression or the top variable (we reuse $V$ as a marker for this).
Our implementation uses a tagged union to define a type that can be either an expression or the top variable.
For the compilation function, we only need to change the definition of $\lazyeta \ z$ and add an extra case for $\lazyeta \ \lambda e$.
Figure~\ref{fig:lazyeta} shows the new compilation function.

Our context merge function $\sqcup$ now also depends on the compiled expression, since we have a special case for $V$.
The updated definition is given in Figure~\ref{fig:lazyeta-ctx}.

The $V$ value effectively encodes an additional implicit used element in the context.
We also have some new cases for the semantic function, available in Figure~\ref{fig:lazyeta-sem}.

@@

\subsection{Linear algorithm}
The implementation of the linear algorithm is much simpler because it is based on the $\strict$ algorithm.
The compilation function is identical save for a different semantic function $\coprod_{\text{linear}}$.
For the full definition, see Figure~\ref{fig:linear}.

In figure~\ref{fig:linear-sem} we give the corresponding semantic function, which is no longer recursive, making this algorithm linear.

@@

\begin{figure}
    \begin{equation*}
        \begin{array}{l l l l}
            \lazy & z         & = & t \models I                                                             \\
            \lazy & s \ e     & = & \Gamma,f \models c                                                      \\
                  &           &   & \text{where } (\Gamma \models c) := \lazy e                             \\
            \\
            \lazy & \lambda e & = & \emptyset \models K \; c                                                \\
                  &           &   & \text{where } (\emptyset \models c) := \lazy e                          \\
            \lazy & \lambda e & = & (\emptyset \models K) \coprod (\Gamma \models c)                        \\
                  &           &   & \text{where } (\Gamma,f \models c) := \lazy e                           \\
            \lazy & \lambda e & = & \Gamma \models K \; c                                                   \\
                  &           &   & \text{where } (\Gamma,t \models c) := \lazy e                           \\
            \\
            \lazy & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)) \\
                  &           &   & \text{where } (\Gamma_1 \models c_1) := \lazy e_1                       \\
                  &           &   & \text{where } (\Gamma_2 \models c_2) := \lazy e_2                       \\
                  &           &   & \text{where } \Gamma := \Gamma_1 \sqcup \Gamma_2                        \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Compilation function with Lazy Weakening.}
    \label{fig:lazy}
\end{figure}

\begin{figure}
    \begin{equation*}
        \begin{array}{l l l l l}
            \Gamma_1    & \sqcup & \emptyset   & = & \Gamma_1                     \\
            \emptyset   & \sqcup & \Gamma_2    & = & \Gamma_2                     \\
            \Gamma_1,t  & \sqcup & \Gamma_2,\_ & = & (\Gamma_1 \sqcup \Gamma_2),t \\
            \Gamma_1,\_ & \sqcup & \Gamma_2,t  & = & (\Gamma_1 \sqcup \Gamma_2),t \\
            \Gamma_1,f  & \sqcup & \Gamma_2,f  & = & (\Gamma_1 \sqcup \Gamma_2),f \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Context merge function for Lazy Weakening.}
    \label{fig:lazy-ctx}
\end{figure}

\begin{figure}
    \begin{equation*}
        \arraycolsep=1.4pt
        \begin{array}{l l l l l}
            (\emptyset \models c_1)  & \coprod & (\emptyset \models c_2)  & = & c_1 \ c_2                                                \\
            \\
            (\emptyset \models c_1)  & \coprod & (\Gamma_2,t \models c_2) & = & (\emptyset \models B c_1) \coprod (\Gamma_2 \models c_2) \\
            \\
            (\Gamma_1,t \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\emptyset \models C C c_2) \coprod (\Gamma \models c_1) \\
            \\
            (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (s \coprod l)) \coprod r)              \\
                                     &         &                          &   & \text{where } s := (\emptyset \models S )                \\
                                     &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                     &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
            \\
            % The new additions
            (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)    \\
            \\
            (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (b \coprod l)) \coprod r)              \\
                                     &         &                          &   & \text{where } b := (\emptyset \models B )                \\
                                     &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                     &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
            \\
            (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models (c \coprod l)) \coprod r)              \\
                                     &         &                          &   & \text{where } c := (\emptyset \models C )                \\
                                     &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                     &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
            \\
            (\emptyset \models c_1)  & \coprod & (\Gamma_2,f \models c_2) & = & (\emptyset \models c_1) \coprod (\Gamma_2 \models c_2)   \\
            \\
            (\Gamma_1,f \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\Gamma_1 \models c_1) \coprod (\emptyset \models c_2)   \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Semantic function with Lazy Weakening.}
    \label{fig:lazy-sem}
\end{figure}

\begin{figure}
    \begin{equation*}
        \begin{array}{l l l l}
            \lazyeta & z         & = & \emptyset \models V                                                          \\
            \lazyeta & s \ e     & = & \Gamma,f \models c                                                           \\
                     &           &   & \text{where } (\Gamma \models c) := \lazyeta e                               \\
            \\
            \lazyeta & \lambda e & = & \emptyset \models I                                                          \\
                     &           &   & \text{where } (\emptyset \models V) := \lazyeta e                            \\
            \lazyeta & \lambda e & = & \emptyset \models K \; c                                                     \\
                     &           &   & \text{where } (\emptyset \models c) := \lazyeta e                            \\
            \lazyeta & \lambda e & = & (\emptyset \models K) \coprod_\eta (\Gamma \models c)                        \\
                     &           &   & \text{where } (\Gamma,f \models c) := \lazyeta e                             \\
            \lazyeta & \lambda e & = & \Gamma \models K \; c                                                        \\
                     &           &   & \text{where } (\Gamma,t \models c) := \lazyeta e                             \\
            \\
            \lazyeta & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod_\eta (\Gamma_2 \models c_2)) \\
                     &           &   & \text{where } (\Gamma_1 \models c_1) := \lazyeta e_1                         \\
                     &           &   & \text{where } (\Gamma_2 \models c_2) := \lazyeta e_2                         \\
                     &           &   & \text{where } \Gamma := \Gamma_1 \sqcup_\eta \Gamma_2                        \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Compilation function with Eta optimization.}
    \label{fig:lazyeta}
\end{figure}

\begin{figure}

    \begin{equation*}
        \begin{array}{l l l l l}
            (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\emptyset \models V)     & = & \Gamma_1,t                        \\
            (\emptyset \models V)     & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & \Gamma_2,t                        \\
            (\emptyset \models c_1)   & \sqcup_\eta & (\emptyset \models V)     & = & t                                 \\
            (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models c_2)   & = & t                                 \\
            (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models V)     & = & \mathbf{impossible}               \\
            \\
            % Original ones
            (\Gamma_1 \models c_1)    & \sqcup_\eta & (\emptyset \models c_2)   & = & \Gamma_1                          \\
            (\emptyset \models c_1)   & \sqcup_\eta & (\Gamma_2 \models c_2)    & = & \Gamma_2                          \\
            (\Gamma_1,t \models c_1)  & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
            (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\Gamma_2,t \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
            (\Gamma_1,f \models c_1)  & \sqcup_\eta & (\Gamma_2,f \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),f \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Context merge function for Eta optimization.}
    \label{fig:lazyeta-ctx}
\end{figure}

\begin{figure}
    \begin{equation*}
        \arraycolsep=1.4pt
        \begin{array}{l l l l l}
            % Added for eta optimization
            (\Gamma_1,f \models c_1) & \coprod_\eta & (\emptyset \models V)    & = & c_1                                                        \\
            (\emptyset \models V)    & \coprod_\eta & (\Gamma_2,f \models c_2) & = & (\emptyset \models CI) \coprod_\eta (\Gamma_2 \models c_2) \\

@@

            (\Gamma_1 \models c_1)   & \coprod_\eta & (\Gamma_2 \models)       & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models)          \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Semantic function with Eta optimization.}
    \label{fig:lazyeta-sem}
\end{figure}

\begin{figure}
    \begin{equation*}
        \begin{array}{l l l l}
            \linear & z         & = & 1 \models I                                                              \\
            \linear & s \ e     & = & n+1 \models (0 \models K) \coprod (n \models c)                          \\
                    &           &   & \text{where } (n \models c) := \linear e                                 \\
            \\
            \linear & \lambda e & = & 0 \models K \; c                                                         \\
                    &           &   & \text{where } (0 \models c) := \linear e                                 \\
            \linear & \lambda e & = & (n-1) \models K \; c                                                     \\
                    &           &   & \text{where } (n \models c) := \linear e                                 \\
            \\
            \linear & e_1 \ e_2 & = & (n \models ((n_1 \models c_1) \coprod_{\text{linear}} (n_2 \models c_2)) \\
                    &           &   & \text{where } (n_1 \models c_1) := \linear e_1                           \\
                    &           &   & \text{where } (n_2 \models c_2) := \linear e_2                           \\
                    &           &   & \text{where } n := \max n_1 \, n_2                                       \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Compilation function for Linear algorithm}
    \label{fig:linear}
\end{figure}

\begin{figure}
    \begin{equation*}
        \arraycolsep=1.4pt
        \begin{array}{l l l l l}
            (0 \models c_1)   & \coprod & (0 \models c_2)   & = & c_1 \ c_2                               \\
            \\
            (0 \models c_1)   & \coprod & (n \models c_2)   & = & B_n \ c_1 \ c_2                         \\
            \\
            (n \models c_1)   & \coprod & (0 \models c_2)   & = & C_n \ c_1 \ c_2                         \\
            \\
            (n \models c_1)   & \coprod & (n \models c_2)   & = & S_n \ c_1 \ c_2                         \\
            \\
            (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & B_{n_2-n_1} \ (S_{n_1} \ c_1) \ c_2     \\
                              &         &                   &   & \text{where } n_1 < n_2                 \\
            (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & C_{n_1-n_2} \ (B_{n_1-n_2} \ c_1) \ c_2 \\
                              &         &                   &   & \text{where } n_1 > n_2                 \\
        \end{array}
    \end{equation*}
    \centering
    \caption{Semantic function for Linear algorithm}
    \label{fig:linear-sem}
\end{figure}
